```{r include = FALSE}

```

# Data

The data that is used is Trust and Quality of Local Governance in European cities, [source](http://data.europa.eu/euodp/en/data/dataset/jrc-coin-regio-trust-and-quality-of-local-governance-in-european-cities-2011-2014). The dataset contains 70 cities in 28 eu-countries. The data is collected by multiple surveys. The publisher of the data is European Union and it was published in 2015. The data was collected 2011-2014. The original dataset contains 53 questionnaire variables, two background variable (country, is city capital) and one variable for id (city). The questionnaire variables are all numeric and have values between 0 and 1.

Because dataset has some problems (especially missing values), too much variables and I would like have more background variables, there is done some data wrangling. The wrangling part is done in seperate script that can be found [here](https://github.com/tvkangas/IODS-final/blob/master/wrangling.R).

After the wrangling, the dataset contains 17 variables. There is 12 varibleas from the surveys, but now most of them contains information from multiple original variables. Those 12 varibles contains information from trust and quality of local governance. There is also two variables describing population of the city: population share comparing to the country and the populations. Then there is id-variables and one that tells if the city is a capital. Variables are described thoroughly in separate file that can be found [here](https://github.com/tvkangas/IODS-final/blob/master/dataInformation.txt). In that text file is also source for populations.

```{r}
#Reading the dataset. After reading citynames are used as rownames so after that running number and citynames can be removed from the dataset
trust_dataTMP <- read.csv("C:/Users/Tuukka/Desktop/IODS-final/trustdata.csv", sep=",", dec=".", header = TRUE)
rownames(trust_dataTMP) <- trust_dataTMP$city
trust_dataTMP <- dplyr::select(trust_dataTMP, -X)
trust_dataTMP <- dplyr::select(trust_dataTMP, -city)
```

