---
title: "IODS course final assignment, data visualization"
author: |
  | Tuukka Kangas
  | tuukka.kangas@helsinki.fi
date: March 5th, 2017
runtime: shiny
output: html_document
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(shiny)
library(ggplot2, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
``` 

This document includes data visualization tool for my final assignment in IODS course. GitHub-reposity for project can be found [here](https://github.com/tvkangas/IODS-final). In GitHub is full description for this project.

Because of the high number of the variable combinatios, I wanted to create an interactive tool for doing scatter plots. For that purpose I need to use _shiny_-package. With this tool user can easily see how cities are located when changing conditions for the plotting.


```{r eruptions, echo=FALSE}

visudata <- read.csv("visual_data.csv", sep=";", dec=".", header = TRUE)

rownames(visudata) <- visudata$city
visudata <- dplyr::select(visudata, -city)
visudata$cluster <- as.factor(visudata$cluster)
visudata$overMillion <- as.factor(visudata$overMillion)
visudata$capital <- as.factor(visudata$overMillion)

muuttujat <- c("Fair compensation in court" = 'faircourt', "Safe at night" = 'safe', "Broke in the home" = 'broke', "Discriminated when looking for a job" = 'workdis', "Trust in different authorities" = 'trust', "Corruption in different authorities" = 'corruption', "Citicens' freedom" = 'freedom', "Giving bribe to the authorities" = 'bribe', "House demolition is unfair" ='houdem', "Small business are fined after illegal actions" = 'busfine', "Openess of local government" = 'govopen', "Lawful, police and court"='lawful')

inputPanel(
  selectInput(inputId ="varx", label = "x-variable",
              choices = muuttujat, selected = 'faircourt'),
  selectInput(inputId ="vary", label = "y-variable",
              choices = muuttujat, selected = 'govopen')
)

regFormulaX <- reactive({
    rtn <- (paste(input$varx))
    rtn
  })

regFormulaY <- reactive({
    rtn <- (paste(input$vary))
    rtn
  })


renderPlot({
  
ggplot(data=visudata, aes_string(x=regFormulaX(), y=regFormulaY())) + 
 geom_point(aes(colour=cluster)) +
 geom_text(aes(label=rownames(visudata), colour = cluster), size = 5) +
 labs(x = regFormulaX(), y=regFormulaY(), title="Scatter plot, visualization of different clusters") +
 theme(text = element_text(size=20))
})


```



